---
title: "Explore Friends Birthday"
---

```{r setup, include=FALSE}
library(ggplot2)
install.packages("lubridate")
library(lubridate)
library(dplyr)
```
## Explore a sample birthday data set and answer following questions:
(1) How many people have the same birthday as me?
(2) Which month contains the most number of birthdays?
(3) How many birthdays are in each month?
(4) Which day of the year has the most number of birthdays?
(5) Do you have at least 365 friends that have birthdays on everyday of the year?

```{r load data}
birthdays <- read.csv('birthdaysExample.csv', header = T, row.names = NULL, as.is=TRUE ,check.names = F)
```
#### convert string to datetime
```{r}
birthdays$bday<-strptime(birthdays$dates, '%m/%d/%y')
```
#### how many people have the same birthday as me? Assume my birthday is December 12th.
```{r}
nrow(subset(birthdays, month(bday)=='12' & day(bday)=='12'))
```
#### I have one friend who has the same birthday as me!

#### Which month contains the most number of birthdays?
```{r}
birthdays$month <-month(birthdays$bday)
birthdays$bday <- as.Date(birthdays$bday)
month_counts <-count(birthdays, month)
month_counts <- arrange(month_counts, desc(n))
```
#### March has most number of birthdays, 98

#### Question (3) How many birthdays are in each month? 
```{r}
ggplot(aes(x=month, y=n), data=month_counts)+
  ggtitle("birthday counts by month")+
  geom_bar(stat="identity", color="gray", fill="blue")+
  scale_x_continuous(breaks=seq(0,12,1))
  xlab('month')+
  ylab("number of birthdays")
```
#### Question (4) Which day of the year has the most number of birthdays?
```{r}
birthdays$day <- day(birthdays$bday)
birthdays$day <- factor(birthdays$day)
day_counts <- count(birthdays, day)
day_counts<-arrange(day_counts, desc(n))
```
#### The 14th day has the most number of birthdays.

#### Question (5) Do you have at least 365 friends that have birthdays on everyday of the year?
```{r}
# count the number of unique values of month-day combination
birthdays$month_day<-format(as.Date(birthdays$bday), "%m-%d" )
monthday_counts <-count(birthdays, month_day)
monthday_counts<-arrange(monthday_counts, desc(n))
length(unique(birthdays$month_day))
```
#### There are 1033 birthdays, larger than 365. But there are only 348 unique month-days, and on somedays I dont have a friend whose birthday is that day.

## Summary:
(1) How many people have the same birthday as me?
1
(2) Which month contains the most number of birthdays?
March
(3) How many birthdays are in each month?
Please see the bar chart above
(4) Which day of the year has the most number of birthdays?
14th, when not considering which month.
when considering month, these days have the most number of birthdays: Feb.06th, May22nd, and July 16th.
(5) Do you have at least 365 friends that have birthdays on everyday of the year?
No I don't