---
title: "Gapminder Two-Variable exploration"
---
#### I will investigate life expectancy and gdp data from gapmider.
```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(gridExtra)
```


#### load data
```{r load data}
le <- read.csv('indicator life_expectancy_at_birth.csv', header = T, row.names = 1, check.names = F)
gdp<-read.csv('indicator gapminder gdp_per_capita_ppp.csv', header = T, row.names =NULL, check.names = F)
```
#### gdp had 262 observations while le has 260
```{r}
# inspect value of "GDP_per_capita" found two rows with no country names
table(gdp$GDP_per_capita)
#remove these 2 rows by subsetting
gdp <- subset(gdp, GDP_per_capita!="")
#rename rows by country names, remove the country column
rownames(gdp) <- gdp[,1]
gdp <- gdp[, -1]
```
#### gdp contains data from 1800 to 2015;  while lefe_expectancy data contain data from 1800 to 2016.
#### create a data frame that contain year 2015 data
```{r}
data2015<-merge(le["2015"], gdp["2015"], by=0)
names(data2015) <- c("country", 'life_expectancy', "gdp")
```
#### view statistics of gdp
```{r}
summary(data2015$gdp)
```
#### view statistics of life expectancy
```{r}
summary(data2015$life_expectancy)
```

#### View histogram and boxplot of gdp at 2015
```{r}
p1<-qplot(gdp, data=data2015, bins=50)+
  xlab("GDP per capita")+
  ggtitle("histogram of GDP at year 2015")
p2<-ggplot(aes(x=1, y=gdp), data=data2015)+geom_boxplot()
grid.arrange(p1,p2,ncol=2)
ggsave(" GDP.png", plot=grid.arrange(p1,p2,ncol=2))
```
#### the shape of the histogram of GDP is left skewed. Majority countries have low GDP. The higher the GDP, the fewer number of countries.
#### By looking at the boxplot, you can see that around 75% of countries GDP is around or less than 25000. There are a few countries whose GDP is much larger than other countries.

#### view histogram and boxplot of life expectancy at year 2015
```{r}
p3<-qplot(life_expectancy, data=data2015, bins=30)+
  xlab("life expectancy (year)")+
  ggtitle("histogram of life expectancy at year 2015")
p4<-ggplot(aes(x=1, y=life_expectancy), data=data2015)+geom_boxplot()
grid.arrange(p3,p4,ncol=2)
ggsave("life expentancy.png", plot=grid.arrange(p3,p4,ncol=2))
```
#### By look at the histogram you can see that life expectancy is right skewed. By looking at the boxplot you can see that at year 2015, 75% of the countries life expectancy is above 68. 

#### create a scatter plot to view relationship between life_expectancy and gdp
```{r}
p5<-ggplot(aes(x=gdp, y=life_expectancy), data=data2015) +
  geom_point(color="red", alpha=0.5)+
  xlab("gdp (per capita)")+
  ylab("life expectancy (years)")+
  ggtitle("life expectancy vs. gdp at 2015")
p5
ggsave("life expectancy vs gdp.png")
```
#### There seem to be an exponential relationship between gdp and life expectancy
#### tranform gdp using log10
```{r}
p6<-p5+coord_trans(x='log10')
p6
ggsave("life expectancy vs log10(gdp).png")
```
#### After tranformation, life expectancy seem to be lieanrly correlated with log10 of gdp

#### correlation with life expectancy and log10 of gdp
```{r}
with(data2015, cor.test(life_expectancy, log10(gdp)))
```
#### correlation coefficient = 0.8. P value is 2.2e-16. 
### conclusion: life expectancy is strongly correlated with GDP
